---
title: "Nutrient Analysis"
author: "Callie Stephenson"
date: "2024-12-03"
output: html_document
editor_options: 
  chunk_output_type: inline
---
# Set-up

load libraries
```{r}
library(dplyr)
library(car)
library(ggplot2)
library(fishualize)
library(corrplot)
library(PerformanceAnalytics)
library(performance)
library(MuMIn)
library(ggeffects)
library(lmerTest)
library(glmnet)
library(ggstats)
```

ggplot theme
```{r}
custom_theme <- theme_minimal(base_size = 10)+
  theme(
    legend.position = "none",
    plot.title = element_text(hjust = 0.5),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 10)
  )
```


load data
```{r}
response_data <- read.csv("data/T1_response_data.csv")
explanatory_variables <- read.csv("data/explanatory_variables.csv")
all_nut <- read.csv("data/All_Nutrients_Processed.csv")
```

model data
```{r}
model_data <- left_join(response_data, explanatory_variables, by = join_by(CowTagID))

PAC_model_data <- model_data %>%
  filter(Species == "Pocillopora acuta", !is.na(Pin_Number)) %>% 
  mutate(pc1 = -1*pc1)

PRU_model_data <- model_data %>%
  filter(Species == "Porites rus", !is.na(Pin_Number)) %>% 
  mutate(pc1 = -1*pc1)%>%
  mutate(Percent_Change = if_else(Placement_Code == "PRU V13 A", NA_real_, Percent_Change))

PRU_model_data_caged_only <- PRU_model_data %>% 
  filter(Cage_Uncaged == "C")
```

# PCA
Remove seep:
```{r}
nut_no_seep <- all_nut[all_nut$CowTagID != 'VSEEP',]
pulse_columns <- names(all_nut)[grepl("Low_Tide_Mean_", names(all_nut)) & 
                            !grepl("Temperature|pH|Ammonia_umol|TA",names(all_nut))]
nut_no_seep_scaled <- nut_no_seep %>%
  mutate(across(all_of(pulse_columns), scale))
```

Look at correlation between variables:
```{r}
pulse_df <- as.data.frame(lapply(nut_no_seep_scaled[, c(pulse_columns)], as.numeric))
cor_matrix <- cor(pulse_df)
corrplot(cor_matrix, method = "number")
chart.Correlation(pulse_df)
```

PCA:
```{r}
pca.data <- na.omit(nut_no_seep_scaled[, c(pulse_columns)]) 
pca = princomp(pca.data, cor=TRUE)
```

When we make a PCA of 
```{r}
names(pca.data)
```

we can explain 72.2% of the variation of these factors with 1 axis. 
```{r}
summary(pca)
loadings(pca)
```

```{r}
biplot(pca)
```

```{r autoplot}
# pca_plot <- autoplot(pca, loadings = TRUE, loadings.colour = '#47c16e',
#                       loadings.label = TRUE, loadings.label.size = 3, loadings.label.colour = '#47c16e') +
#  labs(
#     title = "Principal Component Analysis of Nutrient Data",
#     x = "Principal Component 1 (72.249%)",
#     y = "Principal Component 2 (15.019%)"
#   ) +
#   xlim(-2.5,4)+
# #  scale_x_reverse(-0.5,0.5) +
#   theme_minimal(base_size = 15) +  # Clean theme
#   theme(
#     plot.title = element_text(hjust = 0.5, size = 12, face = "bold"),  # Center title
#     axis.title = element_text(size = 12)
#   )
# 
# pca_plot
#ggsave(pca_plot, filename = "PCA_Plot.png", dpi = 600, width = 7, height = 5)
```

```{r}
scores <- as.data.frame(pca$scores)
loadings <- as.data.frame(pca$loadings[, 1:2])  # First two components

pca_plot <- ggplot(scores, aes(x = -1*(Comp.1), y = Comp.2)) +
  geom_point() +
  geom_segment(data = loadings, aes(x = 0, y = 0, xend = -3*Comp.1, yend = 3*Comp.2),
               arrow = arrow(length = unit(0.2, "cm")), color = '#084439') +
  geom_text(data = loadings, aes(x = -2*Comp.1, y = 3*Comp.2, label = rownames(loadings)),
            color = '#084439', hjust = -0.2) +
  labs(
    title = "Principal Component Analysis of Nutrient Data",
    x = "Principal Component 1",
    y = "Principal Component 2"
  ) +
  theme_minimal(base_size = 15) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 12, face = "bold"),
    axis.title = element_text(size = 12)
  )+
  scale_x_continuous(limits = c(-5, 5))

pca_plot
ggsave(pca_plot, filename = "output/PCA_Plot.png", dpi = 600, width = 7, height = 5)
```

```{r}
nut_no_seep_scaled$pc1 = -1*(pca$scores[,1]) #what's this data's score on pc1 axis
nut_no_seep_scaled$pc2 = (pca$scores[,2]) #what's this data's score on pc1 axis
```

```{r}
nut_no_seep$pc1 = (pca$scores[,1])
```

```{r}
sd_columns <- names(all_nut)[grepl("sd_", names(all_nut)) & 
                              !grepl("Phosphate_umolL|NN_umolL|Ammonia_umolL|NNP_umolL|Silicate_umolL",names(all_nut))]
```

```{r}
explanatory_variables <- nut_no_seep[,c("CowTagID", "pc1", pulse_columns, sd_columns)]
#write.csv(explanatory_variables, "data/explanatory_variables.csv")
```

# Growth

## P. acuta

### plot raw
```{r}
pac_growth_raw <- ggplot(PAC_model_data, aes(x = pc1, y = Percent_Change, color = Genotype, shape = Cage_Uncaged)) +
  geom_point() +
  labs(
    x = "PC1",
    y = "Percent Change in Buoyant Weight",
    title = " "
    #title = expression(italic("Pocillopora acuta") ~ "Change in Buoyant Weight by Nutrient Pulses")
  ) +
  scale_color_fish_d(option = "Coryphaena_hippurus", direction = -1) +
  theme_minimal()+
  theme(legend.position="none",
        plot.background = element_rect(fill = "white")) +
  guides(color = "none", shape = "none") 

pac_growth_raw

ggsave("output/pac_growth_nutrientpca_raw.png",pac_growth_raw, height = 5, width = 5.5)
```

### test for influence of cage
```{r}
ggplot(PAC_model_data, aes(x = Cage_Uncaged, y = Percent_Change, fill = Cage_Uncaged)) +
  geom_boxplot() +
  geom_jitter(aes(shape = Genotype, alpha = 0.8))+
  scale_fill_fish_d(option = "Acanthurus_olivaceus", alpha = 0.8, begin = 0.3, end = 1)

cage_effect_growth <- lm(formula = Percent_Change ~ Cage_Uncaged, data = PAC_model_data)
summary(cage_effect_growth)
check_model(cage_effect_growth)

cage_effect_growth_plot <- ggeffect(cage_effect_growth) %>% 
  plot(show_data = TRUE, alpha=0.01, jitter = TRUE, color = "darkblue")
```

### test for influence of genoytpe
```{r}
genotype_effect_growth <- lm(formula = Percent_Change ~ Genotype, data = PAC_model_data)
summary(genotype_effect_growth)

genotype_effect_growth_plot <- ggeffect(genotype_effect_growth) %>% 
  plot(show_data = TRUE, alpha=0.01, jitter = TRUE, color = "darkblue")
genotype_effect_growth_plot
```

### linear
```{r}
linear_growth_model_pac <- lmerTest::lmer(Percent_Change ~ pc1 +
                                  (1|Genotype), data = PAC_model_data)
#check_model(linear_growth_model_pac)
summary(linear_growth_model_pac)
Anova(linear_growth_model_pac)
```

This model does not support the hypothesis that nutriencts from SGD contribute to the variance in calcification in Pocillopora acuta


### polynomial model: 

```{r}
polynomial_growth_model_pac <- lmer(log10(Percent_Change) ~ poly(pc1, 2) +
                                  (1|Genotype) + (1|Cage_Uncaged), data = PAC_model_data)
#check_model(polynomial_growth_model_pac)
Anova(polynomial_growth_model_pac)
```

This model does not support the hypothesis that nutriencts from SGD contribute to the variance in calcification in Pocillopora acuta

## P. rus

### plot raw
```{r}
pru_growth_raw <- ggplot(PRU_model_data, aes(x = pc1, y = Percent_Change, color = Genotype, shape = Cage_Uncaged)) +
  geom_point() +
  labs(x = "PC1", y = "Percent Change in Buoyant Weight", title = " ") +
  scale_color_fish_d(option = "Coryphaena_hippurus", direction = -1) +
  theme_minimal()+
  theme(legend.position="none",
        plot.background = element_rect(fill = "white")) +
  guides(color = "none", shape = "none") 

pru_growth_raw
ggsave("output/pru_growth_nutrientpca_raw.png",pru_growth_raw, height = 5, width = 5.5)
```

### test for influence of cage
```{r}
ggplot(PRU_model_data, aes(x = Cage_Uncaged, y = Percent_Change, fill = Cage_Uncaged)) +
  geom_boxplot() +
  geom_text(label = PRU_model_data$Pin_Number)+
  geom_jitter(aes(shape = Genotype, alpha = 0.8))+
  scale_fill_fish_d(option = "Acanthurus_olivaceus", alpha = 0.8, begin = 0.3, end = 1)

cage_effect_growth <- lm(formula = Percent_Change ~ Cage_Uncaged, data = PRU_model_data)
summary(cage_effect_growth)
check_model(cage_effect_growth)

ggcoef_model(cage_effect_growth)
```

### test for influence of genoytpe
```{r}
genotype_effect_growth <- lm(formula = Percent_Change ~ Genotype, data = PRU_model_data)
summary(genotype_effect_growth)
ggcoef_model(genotype_effect_growth)
```
### linear
```{r}
linear_growth_model_pru <- lmerTest::lmer(Percent_Change ~ pc1 +
                                  (1|Genotype) + (1|Cage_Uncaged), data = PRU_model_data)
#check_model(linear_growth_model_pru)
Anova(linear_growth_model_pru)
summary(linear_growth_model_pru)
```

We do not find support for pc1 affecting growth of Porites rus in a linear fashion

### polynomial
```{r}
polynomial_growth_model_pru <- lmer(Percent_Change ~ poly(pc1, 2) +
                                  (1|Genotype) + (1|Cage_Uncaged), data = PRU_model_data)
#check_model(polynomial_growth_model_pru)
Anova(polynomial_growth_model_pru)
```
We do not find support for pc1 affecting growth of Porites rus in a polynomial fashion

## Caged P. rus only

I also ran these models on a dataset with only caged corals:

### linear
```{r}
linear_growth_model_pru_c <- lmer(Percent_Change ~ pc1 +
                                  (1|Genotype), data = PRU_model_data_caged_only)
#check_model(linear_growth_model_pru_c)
Anova(linear_growth_model_pru_c)
```

### polynomial
```{r}
polynomial_growth_model_pru_c <- lmer(Percent_Change ~ poly(pc1, 2) +
                                  (1|Genotype), data = PRU_model_data_caged_only)
#check_model(polynomial_growth_model_pru_c)
Anova(polynomial_growth_model_pru_c)
```

These model does not support the hypothesis that nutriencts from SGD contribute to the variance in calcification in Pocillopora acuta

When only looking at caged samples, we still don't find any influence of PC1 on calcification

# Symbiont Count

## Pocillopora acuta

### plot raw
```{r}
pac_symb_raw <- ggplot(PAC_model_data, aes(x = pc1, y = log10(FSC.Events_per_cm_2), color = Genotype, shape = Cage_Uncaged)) +
  geom_point() +
  labs(x = "PC1", 
       y = expression(paste("Endosymbiont density (cells ", 10^6, cm^2, ")")),
      title = " ") +
  scale_color_fish_d(option = "Coryphaena_hippurus", direction = -1) +
  theme_minimal()+
  theme(legend.position="none",
        plot.background = element_rect(fill = "white")) +
  guides(color = "none", shape = "none") 

pac_symb_raw

ggsave("output/pac_symb_nutrientpca_raw.png",pac_symb_raw, height = 5, width = 5.5)
```
### test for influence of cage
```{r}
ggplot(PAC_model_data, aes(x = Cage_Uncaged, y = log10(FSC.Events_per_cm_2), fill = Cage_Uncaged)) +
  geom_boxplot() +
  geom_jitter(aes(shape = Genotype, alpha = 0.8))+
  scale_fill_fish_d(option = "Acanthurus_olivaceus", alpha = 0.8, begin = 0.3, end = 1)

cage_effect_symb <- lm(formula = log10(FSC.Events_per_cm_2) ~ Cage_Uncaged, data = PAC_model_data)
summary(cage_effect_symb)
check_model(cage_effect_symb)
```

### test for influence of genoytpe
```{r}
genotype_effect_symb <- lm(formula = log10(FSC.Events_per_cm_2) ~ Genotype, data = PAC_model_data)
summary(genotype_effect_symb)
ggcoef_model(genotype_effect_symb)
```
### linear model
```{r}
linear_symb_model_pac <- lmer(log10(FSC.Events_per_cm_2) ~ pc1 +
                                  (1|Genotype) + (1|Cage_Uncaged), data = PAC_model_data)
#check_model(linear_symb_model_pac)
Anova(linear_symb_model_pac)
```

This model does not support the hypothesis that nutriencts from SGD contribute to the variance in symbiont density in Pocillopora acuta

### polynomial model

```{r}
polynomial_symb_model_pac <- lmer(log10(FSC.Events_per_cm_2) ~ poly(pc1, 2) +
                                  (1|Genotype) + (1|Cage_Uncaged), data = PAC_model_data)
#check_model(polynomial_symb_model_pac)
Anova(polynomial_symb_model_pac)
```

This model does not support the hypothesis that nutriencts from SGD contribute to the variance in symbiont density in Pocillopora acuta

## Porites rus
```{r}
pru_symb_raw <- ggplot(PRU_model_data, aes(x = pc1, y = log10(FSC.Events_per_cm_2), color = Genotype, shape = Cage_Uncaged)) +
  geom_point() +
  labs(x = "PC1", 
       y = expression(paste("Endosymbiont density (cells ", 10^6, cm^2, ")")),
      title = " ") +
  scale_color_fish_d(option = "Coryphaena_hippurus", direction = -1) +
  theme_minimal()+
  theme(legend.position="none",
        plot.background = element_rect(fill = "white")) +
  guides(color = "none", shape = "none") 

pru_symb_raw

ggsave("output/pru_symb_nutrientpca_raw.png",pru_symb_raw, height = 5, width = 5.5)
```

### test for influence of cage
```{r}
ggplot(PRU_model_data, aes(x = Cage_Uncaged, y = log10(FSC.Events_per_cm_2), fill = Cage_Uncaged)) +
  geom_boxplot() +
  geom_jitter(aes(shape = Genotype, alpha = 0.8))+
  scale_fill_fish_d(option = "Acanthurus_olivaceus", alpha = 0.8, begin = 0.3, end = 1)

cage_effect_symb <- lm(formula = log10(FSC.Events_per_cm_2) ~ Cage_Uncaged, data = PRU_model_data)
summary(cage_effect_symb)
check_model(cage_effect_symb)

ggcoef_model(cage_effect_symb)
```
No cage effect
### test for influence of genoytpe
```{r}
genotype_effect_symb <- lm(formula = log10(FSC.Events_per_cm_2) ~ Genotype, data = PRU_model_data)
summary(genotype_effect_symb)
ggcoef_model(genotype_effect_symb)

genotype_effect_symb_plot <- ggeffect(genotype_effect_symb) %>% 
  plot(show_data = TRUE, alpha=0.01, jitter = TRUE, color = "darkblue")
```
One genotpe is significantly lower than others.

### linear model
```{r}
linear_symb_model_pac <- lmer(log10(FSC.Events_per_cm_2) ~ pc1 +
                                  (1|Genotype) + (1|Cage_Uncaged), data = PAC_model_data)
#check_model(linear_symb_model_pac)
Anova(linear_symb_model_pac)
```
### linear model:

```{r}
linear_symb_model_pru <- lmer(log10(FSC.Events_per_cm_2) ~ pc1 +
                                  (1|Genotype) + (1|Cage_Uncaged), data = PRU_model_data)
#check_model(linear_symb_model_pru)
Anova(linear_symb_model_pru)
```

We do not find support for pc1 affecting symb of Porites rus in a linear fashion

### polynomial model:

```{r}
polynomial_symb_model_pru <- lmer(log10(FSC.Events_per_cm_2) ~ poly(pc1, 2) +
                                  (1|Genotype) + (1|Cage_Uncaged), data = PRU_model_data)
#check_model(polynomial_symb_model_pru)
Anova(polynomial_symb_model_pru)
```

We do not find support for pc1 affecting symb of Porites rus in a polynomial fashion

## Caged P. rus only

### linear

```{r}
linear_symb_model_pru_c <- lmer(log10(FSC.Events_per_cm_2) ~ pc1 +
                                  (1|Genotype), data = PRU_model_data_caged_only)
#check_model(linear_symb_model_pru_c)
Anova(linear_symb_model_pru_c)
```

### polynomial
```{r}
polynomial_symb_model_pru_c <- lmer(log10(FSC.Events_per_cm_2) ~ poly(pc1, 2) +
                                  (1|Genotype), data = PRU_model_data_caged_only)
#check_model(polynomial_symb_model_pru_c)
Anova(polynomial_symb_model_pru_c)
```

This model does not support the hypothesis that nutrients from SGD contribute to the variance in symbiont density in Porites rus
# Chlorophyll

## Pocillopora acuta
### check raw
```{r}
pac_chl_raw <- ggplot(PAC_model_data, aes(x = pc1, y = Chl_ug.cm.2, color = Genotype, shape = Cage_Uncaged)) +
  geom_point() +
  labs(x = "PC1", 
       y = expression(paste("Chlorophyll density (μg ", cm^2, ")")),
      title = " ") +
  scale_color_fish_d(option = "Coryphaena_hippurus", direction = -1) +
  theme_minimal()+
  theme(legend.position="none",
        plot.background = element_rect(fill = "white")) +
  guides(color = "none", shape = "none") +
  scale_y_continuous(limits = c(0,4.5))

pac_chl_raw

ggsave("output/pac_chl_nutrientpca_raw.png",pac_chl_raw, height = 5, width = 5.5)
```

### test for influence of cage
```{r}
ggplot(PAC_model_data, aes(x = Cage_Uncaged, y = Chl_ug.cm.2, fill = Cage_Uncaged)) +
  geom_boxplot() +
  geom_text(label = PAC_model_data$Pin_Number)+
  geom_jitter(aes(shape = Genotype, alpha = 0.8))+
  scale_fill_fish_d(option = "Acanthurus_olivaceus", alpha = 0.8, begin = 0.3, end = 1)

cage_effect_chla <- lm(formula = Chl_ug.cm.2 ~ Cage_Uncaged, data = PAC_model_data)
summary(cage_effect_chla)
check_model(cage_effect_chla)

cage_effect_chla_plot <- ggeffect(cage_effect_chla) %>% 
  plot(show_data = TRUE, alpha=0.01, jitter = TRUE, color = "darkblue")
cage_effect_chla_plot

ggcoef_model(cage_effect_chla)
```
No cage effect

### test for influence of genoytpe
```{r}
genotype_effect_growth <- lm(formula = Chl_ug.cm.2 ~ Genotype, data = PAC_model_data)
summary(genotype_effect_growth)
ggcoef_model(genotype_effect_growth)
```
One genotype (PA 18) faired better than the others

```{r}
linear_chl_model_pac <- lmer(Chl_ug.cm.2 ~ pc1 +
                                  (1|Genotype) + (1|Cage_Uncaged), data = PAC_model_data)
#check_model(linear_growth_model_pac)
Anova(linear_chl_model_pac)
r2_values <- r.squaredGLMM(linear_chl_model_pac)
marginal_r2 <- r2_values[1]  # Marginal R^2
conditional_r2 <- r2_values[2]
```

### Plot significant model
```{r}
pred_chl <- ggpredict(linear_chl_model_pac, terms = "pc1")

#annnotations
anova_results <- car::Anova(linear_chl_model_pac)
p_values <- anova_results$`Pr(>Chisq)`[1]  # Adjust index based on your model's fixed effects
names(p_values) <- rownames(anova_results)[1] 

plot_chl <- ggplot() +
  # Raw data points
  geom_point(data = PAC_model_data, 
             aes(x = pc1, y = Chl_ug.cm.2), color = "grey50", alpha = 0.6) +
  # Prediction line with confidence interval
  geom_line(data = pred_chl, aes(x = x, y = predicted), color = "#21908c") +
  geom_ribbon(data = pred_chl, aes(x = x, ymin = conf.low, ymax = conf.high), 
              fill = "#21908c", alpha = 0.2) +
  scale_y_continuous(limits = c(0,4.5)) +
  labs(
    x = "PC1",
    y = expression(paste("Chlorophyll density (μg ", cm^2, ")"))
  ) +
  custom_theme+
  annotate("text", y = 0.5, x = max(pred_chl$x, na.rm = TRUE), 
         label = paste("p =", round(p_values["pc1"], 6), 
                       "\nMarginal R² =", round(marginal_r2, 3), 
                       "\nConditional R² =", round(conditional_r2, 3)), 
         hjust = 1.1, vjust = 1, size = 3)
```

```{r}
pac_chl_raw_MOD <- ggplot() +
  geom_point(aes(x = pc1, y = Chl_ug.cm.2, color = Genotype, shape = Cage_Uncaged), data = PAC_model_data) +
  labs(x = "PC1", 
       y = expression(paste("Chlorophyll density (μg ", cm^2, ")")),
      title = " ") +
  scale_color_fish_d(option = "Coryphaena_hippurus", direction = -1) +
  theme_minimal()+
  scale_y_continuous(limits = c(0,4.5)) +
  theme(legend.position="none",
        plot.background = element_rect(fill = "white")) +
  guides(color = "none", shape = "none") +
  geom_line(data = pred_chl, aes(x = x, y = predicted), color = "#21908c") +
  geom_ribbon(data = pred_chl, aes(x = x, ymin = conf.low, ymax = conf.high), 
              fill = "#21908c", alpha = 0.2)+
  annotate("text", y = 0.5, x = max(pred_chl$x, na.rm = TRUE), 
         label = paste("p =", round(p_values["pc1"], 6), 
                       "\nMarginal R² =", round(marginal_r2, 3), 
                       "\nConditional R² =", round(conditional_r2, 3)), 
         hjust = 1.1, vjust = 1, size = 3)

ggsave("output/pac_chl_nutrientpca_mod.png",pac_chl_raw_MOD, height = 5, width = 5.5)
```

#### Univariate Models

```{r}
# Define the data and random effects
random_effects <- "(1|Genotype) + (1|Cage_Uncaged)"

# Define the predictors
predictors <- c("Low_Tide_Mean_Salinity", "Low_Tide_Mean_NN_umolL", 
                "Low_Tide_Mean_Phosphate_umolL", "Low_Tide_Mean_Silicate_umolL", 
                "Low_Tide_Mean_NNP_umolL")

# Initialize a results dataframe
results <- data.frame(Variable = character(), 
                      P_Value = numeric(), 
                      Marginal_R2 = numeric(), 
                      Conditional_R2 = numeric(),
                      AIC = numeric(),
                      stringsAsFactors = FALSE)

# Loop through each predictor
for (variable in predictors) {
  # Construct the formula dynamically
  formula <- as.formula(paste("Chl_ug.cm.2 ~", variable, "+", random_effects))
  
  # Fit the model
  model <- lmer(formula, data = PAC_model_data)
  
  # Get the p-value from Anova()
  p_value <- Anova(model)[[3]]
  
  # Get R^2 values
  r2_values <- r.squaredGLMM(model)
  marginal_r2 <- r2_values[1]  # Marginal R^2
  conditional_r2 <- r2_values[2]  # Conditional R^2
  
  # Get AIC values
  aic_value <- AIC(model)
  
  # Append results to the dataframe
  results <- rbind(results, 
                   data.frame(Variable = variable, 
                              P_Value = p_value, 
                              Marginal_R2 = marginal_r2, 
                              Conditional_R2 = conditional_r2,
                              AIC = aic_value))
}

# Display the results
print(results)
```

#### LASSO
P acuta ~ PC 1 Axes
```{r PAC lasso}
# # Remove the Pin_Number and Species columns
# PAC_lasso <- PAC_model_data[,c("Chl_ug.cm.2", "Genotype", "Cage_Uncaged", pulse_columns)] %>% 
#   mutate(Low_Tide_Mean_Salinity = scale(Low_Tide_Mean_Salinity))%>% 
#   mutate(Low_Tide_Mean_Phosphate_umolL = scale(Low_Tide_Mean_Phosphate_umolL))%>% 
#   mutate(Low_Tide_Mean_Silicate_umolL = scale(Low_Tide_Mean_Silicate_umolL)) %>% 
#   mutate(Low_Tide_Mean_NN_umolL = scale(Low_Tide_Mean_NN_umolL)) %>%
#   mutate(Low_Tide_Mean_NNP_umolL = scale(Low_Tide_Mean_NNP_umolL))
# 
# # Convert Genotype and Cage_Uncaged to factors if they are not already
# PAC_lasso$Genotype <- as.factor(PAC_lasso$Genotype)
# PAC_lasso$Cage_Uncaged <- as.factor(PAC_lasso$Cage_Uncaged)
# 
# # Create dummy variables
# genotype_dummies <- model.matrix( ~ Genotype - 1, data = PAC_lasso)
# cage_uncaged_dummies <- model.matrix( ~ Cage_Uncaged - 1, data = PAC_lasso)
# 
# # Combine the dummy variables with the original data frame (excluding the original columns)
# PAC_lasso <- cbind(PAC_lasso, genotype_dummies, cage_uncaged_dummies)
# PAC_lasso <- PAC_lasso[, !names(PAC_lasso) %in% c("Genotype", "Cage_Uncaged")]
# 
# #lasso works on matrix
# PAC_model_data_matrix <- as.matrix(PAC_lasso)
# 
# x_pac_percent_change <- PAC_model_data_matrix[, -1]
# y_pac_percent_change <- PAC_model_data_matrix[, 1]
# lasso_PAC_growth <- cv.glmnet(x = x_pac_percent_change, y = y_pac_percent_change)
# coef(lasso_PAC_growth)
```

#### Plots
```{r}
linear_chl_NN_pac <- lmer(Chl_ug.cm.2 ~ Low_Tide_Mean_NN_umolL +
                                  (1|Genotype) + (1|Cage_Uncaged), data = PAC_model_data)

r2_values <- r.squaredGLMM(linear_chl_NN_pac)
  marginal_r2 <- r2_values[1]  # Marginal R^2
  conditional_r2 <- r2_values[2] 

pred_chl_NN <- ggpredict(linear_chl_NN_pac, terms = "Low_Tide_Mean_NN_umolL")

#annnotations
anova_results <- car::Anova(linear_chl_NN_pac)
p_values <- anova_results$`Pr(>Chisq)`[1]  # Adjust index based on your model's fixed effects
names(p_values) <- rownames(anova_results)[1] 

plot_chl_NN <- ggplot() +
  geom_point(aes(x = Low_Tide_Mean_NN_umolL, y = Chl_ug.cm.2, color = Genotype, shape = Cage_Uncaged), data = PAC_model_data) +
  labs(x = "Low Tide Mean Nitrate + Nitrite (umolL)", 
       y = expression(paste("Chlorophyll density (μg ", cm^2, ")")),
      title = " ") +
  scale_color_fish_d(option = "Coryphaena_hippurus", direction = -1) +
  theme_minimal()+
  scale_y_continuous(limits = c(0,4.5)) +
  theme(legend.position="none",
        plot.background = element_rect(fill = "white")) +
  guides(color = "none", shape = "none") +
  geom_line(data = pred_chl_NN, aes(x = x, y = predicted), color = "#21908c") +
  geom_ribbon(data = pred_chl_NN, aes(x = x, ymin = conf.low, ymax = conf.high),
              fill = "#21908c", alpha = 0.2)+
  annotate("text", y = 0.5, x = max(pred_chl_NN$x, na.rm = TRUE),
         label = paste("p =", round(p_values["Low_Tide_Mean_NN_umolL"], 6),
                       "\nMarginal R² =", round(marginal_r2, 3),
                       "\nConditional R² =", round(conditional_r2, 3)),
         hjust = 1.1, vjust = 1, size = 3)

ggsave("output/pac_chl_NN_mod1.png", plot_chl_NN, height = 5, width = 5.5)
```

```{r}
linear_chl_NNP_pac <- lmer(Chl_ug.cm.2 ~ Low_Tide_Mean_NNP_umolL +
                                  (1|Genotype) + (1|Cage_Uncaged), data = PAC_model_data)

pred_chl_NNP <- ggpredict(linear_chl_NNP_pac, terms = "Low_Tide_Mean_NNP_umolL")
r2_values <- r.squaredGLMM(linear_chl_NNP_pac)
  marginal_r2 <- r2_values[1]  # Marginal R^2
  conditional_r2 <- r2_values[2] 
#annnotations
anova_results <- car::Anova(linear_chl_NNP_pac)
p_values <- anova_results$`Pr(>Chisq)`[1]  # Adjust index based on your model's fixed effects
names(p_values) <- rownames(anova_results)[1] 

plot_chl_NNP <- ggplot() +
  geom_point(aes(x = Low_Tide_Mean_NNP_umolL, y = Chl_ug.cm.2, color = Genotype, shape = Cage_Uncaged), data = PAC_model_data) +
  labs(x = "Low Tide Ratio of Mean Nitrate + Nitrite (umolL) to Phosphate (umolL)", 
       y = expression(paste("Chlorophyll density (μg ", cm^2, ")")),
      title = " ") +
  scale_color_fish_d(option = "Coryphaena_hippurus", direction = -1) +
  theme_minimal()+
  scale_y_continuous(limits = c(0,4.5)) +
  theme(legend.position="none",
        plot.background = element_rect(fill = "white")) +
  guides(color = "none", shape = "none") +
  geom_line(data = pred_chl_NNP, aes(x = x, y = predicted), color = "#21908c") +
  geom_ribbon(data = pred_chl_NNP, aes(x = x, ymin = conf.low, ymax = conf.high), 
              fill = "#21908c", alpha = 0.2)+
  annotate("text", y = 0.5, x = max(pred_chl_NNP$x, na.rm = TRUE), 
         label = paste("p =", round(p_values["Low_Tide_Mean_NNP_umolL"], 6), 
                       "\nMarginal R² =", round(marginal_r2, 3), 
                       "\nConditional R² =", round(conditional_r2, 3)), 
         hjust = 1.1, vjust = 1, size = 3)
ggsave("output/pac_chl_NNP_mod.png", plot_chl_NNP, height = 5, width = 5.5)
```

### Caged only linear
```{r}
linear_chl_model_pac_caged <- lmer(Chl_ug.cm.2 ~ pc1 +
                                  (1|Genotype), data = PAC_model_data %>% filter(Cage_Uncaged == "C"))
#check_model(linear_growth_model_pac)
Anova(linear_chl_model_pac_caged)
r2_values_caged <- r.squaredGLMM(linear_chl_model_pac_caged)
marginal_r2_caged <- r2_values[1]  # Marginal R^2
conditional_r2_caged <- r2_values[2]
```

```{r}
pred_chl <- ggpredict(linear_chl_model_pac_caged, terms = "pc1")

#annnotations
anova_results <- car::Anova(linear_chl_model_pac_caged)
p_values <- anova_results$`Pr(>Chisq)`[1]  # Adjust index based on your model's fixed effects
names(p_values) <- rownames(anova_results)[1] 

plot_chl_caged <- ggplot() +
  # Raw data points
  geom_point(data = PAC_model_data%>% filter(Cage_Uncaged == "C"), 
             aes(x = pc1, y = Chl_ug.cm.2), color = "grey50", alpha = 0.6) +
  # Prediction line with confidence interval
  geom_line(data = pred_chl, aes(x = x, y = predicted), color = "#21908c") +
  geom_ribbon(data = pred_chl, aes(x = x, ymin = conf.low, ymax = conf.high), 
              fill = "#21908c", alpha = 0.2) +
  scale_y_continuous(limits = c(0,4.5)) +
  labs(
    x = "PC1",
    y = expression(paste("Chlorophyll density (μg ", cm^2, ")"))
  ) +
  custom_theme+
  annotate("text", y = 0.5, x = max(pred_chl$x, na.rm = TRUE), 
         label = paste("p =", round(p_values["pc1"], 6), 
                       "\nMarginal R² =", round(marginal_r2_caged, 3), 
                       "\nConditional R² =", round(conditional_r2_caged, 3)), 
         hjust = 1.1, vjust = 1, size = 3)
```

### Polynomial
```{r}
polynomial_chl_model_pac <- lmer(Chl_ug.cm.2 ~ poly(pc1, 2) +
                                  (1|Genotype) + (1|Cage_Uncaged), data = PAC_model_data)
#check_model(polynomial_growth_model_pru)
Anova(polynomial_chl_model_pac)
```

```{r}
pred_chl_poly <- ggpredict(polynomial_chl_model_pac, terms = "pc1")

#annnotations
anova_results <- car::Anova(polynomial_chl_model_pac)
p_values <- anova_results$`Pr(>Chisq)`[1]  # Adjust index based on your model's fixed effects
names(p_values) <- rownames(anova_results)[1] 

plot_chl <- ggplot() +
  # Raw data points
  geom_point(data = PAC_model_data, 
             aes(x = pc1, y = Chl_ug.cm.2), color = "grey50", alpha = 0.6) +
  # Prediction line with confidence interval
  geom_line(data = pred_chl, aes(x = x, y = predicted), color = "#21908c") +
  geom_ribbon(data = pred_chl, aes(x = x, ymin = conf.low, ymax = conf.high), 
              fill = "#21908c", alpha = 0.2) +
  scale_y_continuous(limits = c(-1.5,25)) +
  labs(
    x = "PC1",
    y = expression(paste("Chlorophyll density (μg ", cm^2, ")"))
  ) +
  custom_theme+
  annotate("text", y = 1, x = max(pred_chl$x, na.rm=TRUE), 
           label = paste("p =", round(p_values["poly(pc1, 2)"], 6)), 
           hjust = 1.1, vjust = 1, size = 3)
```

```{r}
linear_chl_pac_growth <- lmer(Percent_Change ~ Chl_ug.cm.2 +
                                  (1|Genotype) + (1|Cage_Uncaged), data = PAC_model_data)

plot_chl_growth <- ggplot() +
  # Raw data points
  geom_point(data = PAC_model_data, 
             aes(x = Chl_ug.cm.2, y = Percent_Change,
                 color= Genotype, shape = Cage_Uncaged)) +
  labs(
    y = "Percent Change in Buoyant Weight",
    x = expression(paste("Chlorophyll density (μg ", cm^2, ")"))
  ) +
  scale_color_fish_d(option = "Coryphaena_hippurus", direction = -1)+
  custom_theme+
  annotate("text", y = 1, x = 1, 
           label = "p =0.1516", 
           hjust = 1.1, vjust = 1, size = 3)



```


## Porites rus
```{r}
pru_chl_raw <- ggplot(PRU_model_data, aes(x = pc1, y = Chl_ug.cm.2, color = Genotype, shape = Cage_Uncaged)) +
  geom_point() +
  labs(x = "PC1", 
       y = expression(paste("Chlorophyll density (μg ", cm^2, ")")),
      title = " ") +
  scale_color_fish_d(option = "Coryphaena_hippurus", direction = -1) +
  theme_minimal()+
  theme(legend.position="none",
        plot.background = element_rect(fill = "white")) +
  guides(color = "none", shape = "none") 

pru_chl_raw

ggsave("output/pru_chl_nutrientpca_raw.png",pru_chl_raw, height = 5, width = 5.5)
```

### test for influence of cage
```{r}
ggplot(PRU_model_data, aes(x = Cage_Uncaged, y = Chl_ug.cm.2, fill = Cage_Uncaged)) +
  geom_boxplot() +
  geom_text(label = PRU_model_data$Pin_Number)+
  geom_jitter(aes(shape = Genotype, alpha = 0.8))+
  scale_fill_fish_d(option = "Acanthurus_olivaceus", alpha = 0.8, begin = 0.3, end = 1)

cage_effect_chla <- lm(formula = Chl_ug.cm.2 ~ Cage_Uncaged, data = PRU_model_data)
summary(cage_effect_chla)
check_model(cage_effect_chla)

cage_effect_chla_plot <- ggeffect(cage_effect_chla) %>% 
  plot(show_data = TRUE, alpha=0.01, jitter = TRUE, color = "darkblue")
cage_effect_chla_plot

ggcoef_model(cage_effect_chla)
```
I can't tell you why, but there is a cage effect here...

### test for influence of genoytpe
```{r}
genotype_effect_growth <- lm(formula = Chl_ug.cm.2 ~ Genotype, data = PRU_model_data)
summary(genotype_effect_growth)
ggcoef_model(genotype_effect_growth)
```
### linear
```{r}
linear_chl_model_pru <- lmer(Chl_ug.cm.2 ~ pc1 +
                                  (1|Genotype) + (1|Cage_Uncaged), data = PRU_model_data)
#check_model(linear_growth_model_pru)
Anova(linear_chl_model_pru)
```

### polynomial
```{r}
polynomial_chl_model_pru <- lmer(Chl_ug.cm.2 ~ poly(pc1, 2) +
                                  (1|Genotype) + (1|Cage_Uncaged), data = PRU_model_data)
#check_model(polynomial_growth_model_pru)
Anova(polynomial_chl_model_pru)
```

# Heterotrophy
## Carbon

### Pocillopora acuta

#### Exploratory plot:
```{r}
SI_PAC_model_data <- PAC_model_data %>% 
  filter(!is.na(Δ13C))
ggplot(data = SI_PAC_model_data, (aes(y=Δ13C, x= pc1))) +
         geom_point()

pac_13C_raw <- ggplot(PAC_model_data %>% 
  filter(!is.na(Δ13C)), aes(x = pc1, y = Δ13C, color = Genotype)) +
  geom_point() +
  labs(x = "PC1", 
       y = "δ13C",
      title = " ") +
  scale_color_fish_d(option = "Coryphaena_hippurus", direction = -1) +
  theme_minimal()+
  theme(legend.position="none",
        plot.background = element_rect(fill = "white")) +
  guides(color = "none", shape = "none") 

pac_13C_raw

ggsave("output/pac_13C_nutrientpca_raw.png",pac_13C_raw, height = 5, width = 5.5)
```

#### Linear model:
```{r}
linear_carbon_model_pac <- lmer(Δ13C ~ pc1 +
                                  (1|Genotype), data = SI_PAC_model_data)
#check_model(linear_carbon_model_pac)
#summary(linear_carbon_model_pac)
Anova(linear_carbon_model_pac)
```
Not significant

#### Polynomial model:
```{r}
polynomial_carbon_model_pac <- lmer(Δ13C ~ poly(pc1, 2) +
                                  (1|Genotype), data = SI_PAC_model_data)
#check_model(polynomial_carbon_model_pac)
Anova(polynomial_carbon_model_pac)
```
Not significant

### Porites rus
#### Exploratory plot:
```{r}
SI_PRU_model_data <- PRU_model_data %>% 
  filter(!is.na(Δ13C))
ggplot(data = SI_PRU_model_data, (aes(y=Δ13C, x= pc1))) +
         geom_point()

pru_13C_raw <- ggplot(PRU_model_data %>% 
  filter(!is.na(Δ13C)), aes(x = pc1, y = Δ13C, color = Genotype)) +
  geom_point() +
  labs(x = "PC1", 
       y = "δ13C",
      title = " ") +
  scale_color_fish_d(option = "Coryphaena_hippurus", direction = -1) +
  theme_minimal()+
  theme(legend.position="none",
        plot.background = element_rect(fill = "white")) +
  guides(color = "none", shape = "none") 

pru_13C_raw

ggsave("output/pru_13C_nutrientpca_raw.png",pru_13C_raw, height = 5, width = 5.5)
```

#### Linear model:
```{r}
linear_carbon_model_PRU <- lmer(Δ13C ~ pc1 +
                                  (1|Genotype), data = SI_PRU_model_data)
#check_model(linear_carbon_model_PRU)
#summary(linear_carbon_model_PRU)
Anova(linear_carbon_model_PRU)
```
Not significant

#### Polynomial model:
```{r}
polynomial_carbon_model_PRU <- lmer(Δ13C ~ poly(pc1, 2) +
                                  (1|Genotype), data = SI_PRU_model_data)
#check_model(polynomial_carbon_model_PRU)
Anova(polynomial_carbon_model_PRU)
```
Not significant

## Nitrogen

### Pocillopora acuta

#### Exploratory plot:
```{r}
SI_PAC_model_data <- PAC_model_data %>% 
  filter(!is.na(Δ15N))
ggplot(data = SI_PAC_model_data, (aes(y=Δ15N, x= pc1))) +
         geom_point()

pac_15N_raw <- ggplot(PAC_model_data %>% 
  filter(!is.na(Δ13C)), aes(x = pc1, y = Δ13C, color = Genotype)) +
  geom_point() +
  labs(x = "PC1", 
       y = "δ15N",
      title = " ") +
  scale_color_fish_d(option = "Coryphaena_hippurus", direction = -1) +
  theme_minimal()+
  theme(legend.position="none",
        plot.background = element_rect(fill = "white")) +
  guides(color = "none", shape = "none") 

pac_15N_raw

ggsave("output/pac_15N_nutrientpca_raw.png",pac_15N_raw, height = 5, width = 5.5)
```

#### Linear model:
```{r}
linear_nitrogen_model_pac <- lmer(Δ15N ~ pc1 +
                                  (1|Genotype), data = SI_PAC_model_data)
#check_model(linear_nitrogen_model_pac)
#summary(linear_nitrogen_model_pac)
Anova(linear_nitrogen_model_pac)
```
HIGHLY Significant!!!

```{r}
prediction_data <- data.frame(pc1 = seq(min(SI_PAC_model_data$pc1), 
                                   max(SI_PAC_model_data$pc1), 
                                   length.out = 100))
# Add predicted values to the dataframe using the model without random effects
predictions <- predict(linear_nitrogen_model_pac, newdata = prediction_data, re.form = NA, se.fit = TRUE)

#Add to new_data
prediction_data <- prediction_data %>%
  mutate(fit = predictions$fit,
         se.fit = predictions$se.fit,
         lower = fit - 1.96 * se.fit,  # 95% CI lower
         upper = fit + 1.96 * se.fit)   # 95% CI upper

# Create the plot with the fitted line
ggplot() +
  geom_point(data = SI_PAC_model_data, aes(x = pc1, y = Δ15N), color = "grey") +
  geom_line(data = prediction_data, aes(x = pc1, y = fit), color = "#21908c") +  # Add the fitted line
  geom_ribbon(data = prediction_data, aes(x = pc1, ymin = lower, ymax = upper), alpha = 0.2, fill = "#21908c") +  # Add CI
  labs(title = "Porites rus Δ15N (host minus symbiont, proxy for heterotrophy)",
       x = "PC1",
       y = "Δ15N") +
  theme_minimal()
```


#### Polynomial model:
```{r}
polynomial_nitrogen_model_pac <- lmer(Δ15N ~ poly(pc1, 2) +
                                  (1|Genotype), data = SI_PAC_model_data)
#check_model(polynomial_nitrogen_model_pac)
Anova(polynomial_nitrogen_model_pac)
```
Significant

```{r}
prediction_data <- data.frame(pc1 = seq(min(SI_PAC_model_data$pc1), 
                                   max(SI_PAC_model_data$pc1), 
                                   length.out = 100))
# Add predicted values to the dataframe using the model without random effects
predictions <- predict(polynomial_nitrogen_model_pac, newdata = prediction_data, re.form = NA, se.fit = TRUE)

#Add to new_data
prediction_data <- prediction_data %>%
  mutate(fit = predictions$fit,
         se.fit = predictions$se.fit,
         lower = fit - 1.96 * se.fit,  # 95% CI lower
         upper = fit + 1.96 * se.fit)   # 95% CI upper

# Create the plot with the fitted line
ggplot() +
  geom_point(data = SI_PAC_model_data, aes(x = pc1, y = Δ15N), color = "grey") +
  geom_line(data = prediction_data, aes(x = pc1, y = fit), color = "#21908c") +  # Add the fitted line
  geom_ribbon(data = prediction_data, aes(x = pc1, ymin = lower, ymax = upper), alpha = 0.2, fill = "#21908c") +  # Add CI
  labs(title = "Porites rus Δ15N (host minus symbiont, proxy for heterotrophy)",
       x = "PC1",
       y = "Δ15N") +
  theme_minimal()
```

### Porites rus
#### Exploratory plot:
```{r}
SI_PRU_model_data <- PRU_model_data %>% 
  filter(!is.na(Δ15N))
ggplot(data = SI_PRU_model_data, (aes(y=Δ15N, x= pc1))) +
         geom_point()

pru_15N_raw <- ggplot(PRU_model_data %>% 
  filter(!is.na(Δ13C)), aes(x = pc1, y = Δ13C, color = Genotype)) +
  geom_point() +
  labs(x = "PC1", 
       y = "δ15N",
      title = " ") +
  scale_color_fish_d(option = "Coryphaena_hippurus", direction = -1) +
  theme_minimal()+
  theme(legend.position="none",
        plot.background = element_rect(fill = "white")) +
  guides(color = "none", shape = "none") 

pru_15N_raw

ggsave("output/pru_15N_nutrientpca_raw.png",pru_15N_raw, height = 5, width = 5.5)
```

#### Linear model:
```{r}
linear_nitrogen_model_PRU <- lmer(Δ15N ~ pc1 +
                                  (1|Genotype), data = SI_PRU_model_data)
#check_model(linear_nitrogen_model_PRU)
#summary(linear_nitrogen_model_PRU)
Anova(linear_nitrogen_model_PRU)
```
Not significant

#### Polynomial model:
```{r}
polynomial_nitrogen_model_PRU <- lmer(Δ15N ~ poly(pc1, 2) +
                                  (1|Genotype), data = SI_PRU_model_data)
#check_model(polynomial_nitrogen_model_PRU)
Anova(polynomial_nitrogen_model_PRU)
```
Significant

```{r}
prediction_data <- data.frame(pc1 = seq(min(SI_PRU_model_data$pc1), 
                                   max(SI_PRU_model_data$pc1), 
                                   length.out = 100))
# Add predicted values to the dataframe using the model without random effects
predictions <- predict(polynomial_nitrogen_model_PRU, newdata = prediction_data, re.form = NA, se.fit = TRUE)

#Add to new_data
prediction_data <- prediction_data %>%
  mutate(fit = predictions$fit,
         se.fit = predictions$se.fit,
         lower = fit - 1.96 * se.fit,  # 95% CI lower
         upper = fit + 1.96 * se.fit)   # 95% CI upper

# Create the plot with the fitted line
ggplot() +
  geom_point(data = SI_PRU_model_data, aes(x = pc1, y = Δ15N), color = "grey") +
  geom_line(data = prediction_data, aes(x = pc1, y = fit), color = "#21908c") +  # Add the fitted line
  geom_ribbon(data = prediction_data, aes(x = pc1, ymin = lower, ymax = upper), alpha = 0.2, fill = "#21908c") +  # Add CI
  labs(title = "Porites rus Δ15N (host minus symbiont, proxy for heterotrophy)",
       x = "PC1",
       y = "Δ15N") +
  theme_minimal()
```


```{r}
PAC_model_data
```

