---
title: "Caged_Nutrient_Analysis"
author: "Callie Stephenson"
date: "2025-01-03"
output: html_document
editor_options: 
  chunk_output_type: inline
---
# Set-up

load libraries
```{r}
library(dplyr)
library(car)
library(ggplot2)
library(fishualize)
library(corrplot)
library(PerformanceAnalytics)
library(performance)
library(MuMIn)
library(ggeffects)
library(lmerTest)
library(glmnet)
```

ggplot theme
```{r}
custom_theme <- theme_minimal(base_size = 10)+
  theme(
    legend.position = "none",
    plot.title = element_text(hjust = 0.5),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 10)
  )
```


load data
```{r}
response_data <- read.csv("data/T1_response_data.csv")
explanatory_variables <- read.csv("data/explanatory_variables.csv")
all_nut <- read.csv("data/All_Nutrients_Processed.csv")
```

model data
```{r}
model_data <- left_join(response_data, explanatory_variables, by = join_by(CowTagID))

PAC_model_data <- model_data %>%
  filter(Species == "Pocillopora acuta", !is.na(Pin_Number)) %>% 
  mutate(pc1 = -1*pc1)%>% 
  filter(Cage_Uncaged == "C")

PRU_model_data <- model_data %>%
  filter(Species == "Porites rus", !is.na(Pin_Number)) %>% 
  mutate(pc1 = -1*pc1) %>% 
  filter(Cage_Uncaged == "C")
```

# PCA
Remove seep:
```{r}
nut_no_seep <- all_nut[all_nut$CowTagID != 'VSEEP',]
pulse_columns <- names(all_nut)[grepl("Low_Tide_Mean_", names(all_nut)) & 
                            !grepl("Temperature|Ammonia_umol|NNP",names(all_nut))] #You took out NNP  
nut_no_seep_scaled <- nut_no_seep %>%
  mutate(across(all_of(pulse_columns), scale))
```

Look at correlation between variables:
```{r}
pulse_df <- as.data.frame(lapply(nut_no_seep_scaled[, c(pulse_columns)], as.numeric))
cor_matrix <- cor(pulse_df)
corrplot(cor_matrix, method = "number")
chart.Correlation(pulse_df)
```

PCA:
```{r}
pca.data <- na.omit(nut_no_seep_scaled[, c(pulse_columns)]) 
pca = princomp(pca.data, cor=TRUE)
```

When we make a PCA of 
```{r}
names(pca.data)
```

we can explain 73.4% of the variation of these factors with 1 axis. 
```{r}
summary(pca)
loadings(pca)
```

```{r}
biplot(pca)
```

```{r}
# nut_no_seep_scaled$pc1 = -1*(pca$scores[,1]) #what's this data's score on pc1 axis
# nut_no_seep_scaled$pc2 = (pca$scores[,2]) #what's this data's score on pc1 axis
```

```{r}
nut_no_seep$pc1 = -1*(pca$scores[,1]) #changed the orientation but kept it in
```

```{r}
explanatory_variables <- nut_no_seep[,c("CowTagID", "pc1", pulse_columns)]
#write.csv(explanatory_variables, "data/explanatory_variables.csv")
```

# Growth

## P. acuta

### plot raw
```{r}
pac_growth_raw <- ggplot(PAC_model_data, aes(x = pc1, y = Percent_Change, color = Genotype)) +
  geom_point() +
  labs(
    x = "PC1",
    y = "Percent Change in Buoyant Weight",
    title = " "
    #title = expression(italic("Pocillopora acuta") ~ "Change in Buoyant Weight by Nutrient Pulses")
  ) +
  scale_color_fish_d(option = "Coryphaena_hippurus", direction = -1) +
  theme_minimal()+
  theme(legend.position="none",
        plot.background = element_rect(fill = "white")) +
  guides(color = "none", shape = "none") 

pac_growth_raw

#ggsave("output/caged_pac_growth_nutrientpca_raw.png",pac_growth_raw, height = 5, width = 5.5)
```

### linear
```{r}
linear_growth_model_pac <- lmerTest::lmer(Percent_Change ~ pc1 +
                                  (1|Genotype), data = PAC_model_data)
#check_model(linear_growth_model_pac)
#summary(linear_growth_model_pac)
Anova(linear_growth_model_pac)
```

This model does not support the hypothesis that nutriencts from SGD contribute to the variance in calcification in Pocillopora acuta


### polynomial model: 

```{r}
polynomial_growth_model_pac <- lmer(log10(Percent_Change) ~ poly(pc1, 2) +
                                  (1|Genotype), data = PAC_model_data)
#check_model(polynomial_growth_model_pac)
Anova(polynomial_growth_model_pac)
```

This model does not support the hypothesis that nutriencts from SGD contribute to the variance in calcification in Pocillopora acuta

## P. rus

### plot raw
```{r}
pru_growth_raw <- ggplot(PRU_model_data, aes(x = pc1, y = Percent_Change, color = Genotype)) +
  geom_point() +
  labs(x = "PC1", y = "Percent Change in Buoyant Weight", title = " ") +
  scale_color_fish_d(option = "Coryphaena_hippurus", direction = -1) +
  theme_minimal()+
  theme(legend.position="none",
        plot.background = element_rect(fill = "white")) +
  guides(color = "none", shape = "none") 

pru_growth_raw
#ggsave("output/caged_pru_growth_nutrientpca_raw.png",pru_growth_raw, height = 5, width = 5.5)
```

### linear
```{r}
linear_growth_model_pru <- lmerTest::lmer(Percent_Change ~ pc1 +
                                  (1|Genotype), data = PRU_model_data)
#check_model(linear_growth_model_pru)
Anova(linear_growth_model_pru)
#summary(linear_growth_model_pru)
```

We do not find support for pc1 affecting growth of Porites rus in a linear fashion

### polynomial
```{r}
polynomial_growth_model_pru <- lmer(Percent_Change ~ poly(pc1, 2) +
                                  (1|Genotype), data = PRU_model_data)
#check_model(polynomial_growth_model_pru)
Anova(polynomial_growth_model_pru)
```
We do not find support for pc1 affecting growth of Porites rus in a polynomial fashion

# Symbiont Count

## Pocillopora acuta

```{r}
pac_symb_raw <- ggplot(PAC_model_data, aes(x = pc1, y = log10(FSC.Events_per_cm_2), color = Genotype, shape = Cage_Uncaged)) +
  geom_point() +
  labs(x = "PC1", 
       y = expression(paste("Endosymbiont density (cells ", 10^6, cm^2, ")")),
      title = " ") +
  scale_color_fish_d(option = "Coryphaena_hippurus", direction = -1) +
  theme_minimal()+
  theme(legend.position="none",
        plot.background = element_rect(fill = "white")) +
  guides(color = "none", shape = "none") 

pac_symb_raw

#ggsave("output/pac_symb_nutrientpca_raw.png",pac_symb_raw, height = 5, width = 5.5)
```

### linear model
```{r}
linear_symb_model_pac <- lmer(log10(FSC.Events_per_cm_2) ~ pc1 +
                                  (1|Genotype), data = PAC_model_data)
#check_model(linear_symb_model_pac)
Anova(linear_symb_model_pac)
```

This model does not support the hypothesis that nutriencts from SGD contribute to the variance in symbiont density in Pocillopora acuta

### polynomial model

```{r}
polynomial_symb_model_pac <- lmer(log10(FSC.Events_per_cm_2) ~ poly(pc1, 2) +
                                  (1|Genotype), data = PAC_model_data)
#check_model(polynomial_symb_model_pac)
Anova(polynomial_symb_model_pac)
```

This model does not support the hypothesis that nutriencts from SGD contribute to the variance in symbiont density in Pocillopora acuta

## Porites rus
```{r}
pru_symb_raw <- ggplot(PRU_model_data, aes(x = pc1, y = log10(FSC.Events_per_cm_2), color = Genotype, shape = Cage_Uncaged)) +
  geom_point() +
  labs(x = "PC1", 
       y = expression(paste("Endosymbiont density (cells ", 10^6, cm^2, ")")),
      title = " ") +
  scale_color_fish_d(option = "Coryphaena_hippurus", direction = -1) +
  theme_minimal()+
  theme(legend.position="none",
        plot.background = element_rect(fill = "white")) +
  guides(color = "none", shape = "none") 

pru_symb_raw

#ggsave("output/pru_symb_nutrientpca_raw.png",pru_symb_raw, height = 5, width = 5.5)
```

### linear model:

```{r}
linear_symb_model_pru <- lmer(log10(FSC.Events_per_cm_2) ~ pc1 +
                                  (1|Genotype), data = PRU_model_data)
#check_model(linear_symb_model_pru)
Anova(linear_symb_model_pru)
```

We do not find support for pc1 affecting symb of Porites rus in a linear fashion

### polynomial model:

```{r}
polynomial_symb_model_pru <- lmer(log10(FSC.Events_per_cm_2) ~ poly(pc1, 2) +
                                  (1|Genotype), data = PRU_model_data)
#check_model(polynomial_symb_model_pru)
Anova(polynomial_symb_model_pru)
```

We do not find support for pc1 affecting symb of Porites rus in a polynomial fashion

## Caged P. rus only

### linear

```{r}
linear_symb_model_pru_c <- lmer(log10(FSC.Events_per_cm_2) ~ pc1 +
                                  (1|Genotype), data = PRU_model_data_caged_only)
#check_model(linear_symb_model_pru_c)
Anova(linear_symb_model_pru_c)
```

### polynomial
```{r}
polynomial_symb_model_pru_c <- lmer(log10(FSC.Events_per_cm_2) ~ poly(pc1, 2) +
                                  (1|Genotype), data = PRU_model_data_caged_only)
#check_model(polynomial_symb_model_pru_c)
Anova(polynomial_symb_model_pru_c)
```

This model does not support the hypothesis that nutrients from SGD contribute to the variance in symbiont density in Porites rus
# Chlorophyll

## Pocillopora acuta
```{r}
pac_chl_raw <- ggplot(PAC_model_data, aes(x = pc1, y = Chl_ug.cm.2, color = Genotype, shape = Cage_Uncaged)) +
  geom_point() +
  labs(x = "PC1", 
       y = expression(paste("Chlorophyll density (μg ", cm^2, ")")),
      title = " ") +
  scale_color_fish_d(option = "Coryphaena_hippurus", direction = -1) +
  theme_minimal()+
  theme(legend.position="none",
        plot.background = element_rect(fill = "white")) +
  guides(color = "none", shape = "none") +
  scale_y_continuous(limits = c(0,4.5))

pac_chl_raw

#ggsave("output/pac_chl_nutrientpca_raw.png",pac_chl_raw, height = 5, width = 5.5)
```

```{r}
linear_chl_model_pac <- lmer(Chl_ug.cm.2 ~ pc1 +
                                  (1|Genotype), data = PAC_model_data)
#check_model(linear_growth_model_pac)
Anova(linear_chl_model_pac)
r2_values <- r.squaredGLMM(linear_chl_model_pac)
marginal_r2 <- r2_values[1]  # Marginal R^2
conditional_r2 <- r2_values[2]
```

### Plot significant model
```{r}
range_pc1 <- range(PAC_model_data$pc1, na.rm = TRUE)
pred_chl <- ggpredict(linear_chl_model_pac, terms = "pc1")
pred_chl_filtered <- pred_chl %>%
  filter(x >= range_pc1[1] & x <= range_pc1[2])

#annnotations
anova_results <- car::Anova(linear_chl_model_pac)
p_values <- anova_results$`Pr(>Chisq)`[1]  # Adjust index based on your model's fixed effects
names(p_values) <- rownames(anova_results)[1] 

plot_chl <- ggplot() +
  # Raw data points
  geom_point(data = PAC_model_data, 
             aes(x = pc1, y = Chl_ug.cm.2), color = "grey50", alpha = 0.6) +
  # Prediction line with confidence interval
  geom_line(data = pred_chl_filtered, aes(x = x, y = predicted), color = "#21908c") +
  geom_ribbon(data = pred_chl_filtered, aes(x = x, ymin = conf.low, ymax = conf.high), 
              fill = "#21908c", alpha = 0.2) +
  scale_y_continuous(limits = c(0,4.5)) +
  labs(
    x = "PC1",
    y = expression(paste("Chlorophyll density (μg ", cm^2, ")"))
  ) +
  custom_theme+
  annotate("text", y = 0.5, x = max(pred_chl$x, na.rm = TRUE), 
         label = paste("p =", round(p_values["pc1"], 6), 
                       "\nMarginal R² =", round(marginal_r2, 3), 
                       "\nConditional R² =", round(conditional_r2, 3)), 
         hjust = 1.1, vjust = 1, size = 3)

plot_chl
```

```{r}
pac_chl_MOD <- ggplot() +
  geom_point(aes(x = pc1, y = Chl_ug.cm.2, color = Genotype, shape = Cage_Uncaged), data = PAC_model_data) +
  labs(x = "PC1", 
       y = expression(paste("Chlorophyll density (μg ", cm^2, ")")),
      title = " ") +
  scale_color_fish_d(option = "Coryphaena_hippurus", direction = -1) +
  theme_minimal()+
  scale_y_continuous(limits = c(0,4.5)) +
  theme(legend.position="none",
        plot.background = element_rect(fill = "white")) +
  guides(color = "none", shape = "none") +
  geom_line(data = pred_chl, aes(x = x, y = predicted), color = "#21908c") +
  geom_ribbon(data = pred_chl, aes(x = x, ymin = conf.low, ymax = conf.high), 
              fill = "#21908c", alpha = 0.2)+
  annotate("text", y = 0.5, x = max(pred_chl$x, na.rm = TRUE), 
         label = paste("p =", round(p_values["pc1"], 6), 
                       "\nMarginal R² =", round(marginal_r2, 3), 
                       "\nConditional R² =", round(conditional_r2, 3)), 
         hjust = 1.1, vjust = 1, size = 3)

pac_chl_raw_MOD

#ggsave("output/pac_chl_nutrientpca_mod.png",pac_chl_raw_MOD, height = 5, width = 5.5)
```

#### Univariate Models

```{r}
# Define the data and random effects
random_effects <- "(1|Genotype)"

# Define the predictors
predictors <- c("Low_Tide_Mean_Salinity", "Low_Tide_Mean_NN_umolL", 
                "Low_Tide_Mean_Phosphate_umolL", "Low_Tide_Mean_Silicate_umolL", 
                "Low_Tide_Mean_NNP_umolL")

# Initialize a results dataframe
results <- data.frame(Variable = character(), 
                      P_Value = numeric(), 
                      Marginal_R2 = numeric(), 
                      Conditional_R2 = numeric(),
                      AIC = numeric(),
                      stringsAsFactors = FALSE)

# Loop through each predictor
for (variable in predictors) {
  # Construct the formula dynamically
  formula <- as.formula(paste("Chl_ug.cm.2 ~", variable, "+", random_effects))
  
  # Fit the model
  model <- lmer(formula, data = PAC_model_data)
  
  # Get the p-value from Anova()
  p_value <- Anova(model)[[3]]
  
  # Get R^2 values
  r2_values <- r.squaredGLMM(model)
  marginal_r2 <- r2_values[1]  # Marginal R^2
  conditional_r2 <- r2_values[2]  # Conditional R^2
  
  # Get AIC values
  aic_value <- AIC(model)
  
  # Append results to the dataframe
  results <- rbind(results, 
                   data.frame(Variable = variable, 
                              P_Value = p_value, 
                              Marginal_R2 = marginal_r2, 
                              Conditional_R2 = conditional_r2,
                              AIC = aic_value))
}

# Display the results
print(results)
```


#### Plots
```{r}
linear_chl_NN_pac <- lmer(Chl_ug.cm.2 ~ Low_Tide_Mean_NN_umolL +
                                  (1|Genotype), data = PAC_model_data)

r2_values <- r.squaredGLMM(linear_chl_NN_pac)
  marginal_r2 <- r2_values[1]  # Marginal R^2
  conditional_r2 <- r2_values[2] 

pred_chl_NN <- ggpredict(linear_chl_NN_pac, terms = "Low_Tide_Mean_NN_umolL")

#annnotations
anova_results <- car::Anova(linear_chl_NN_pac)
p_values <- anova_results$`Pr(>Chisq)`[1]  # Adjust index based on your model's fixed effects
names(p_values) <- rownames(anova_results)[1] 

plot_chl_NN <- ggplot() +
  geom_point(aes(x = Low_Tide_Mean_NN_umolL, y = Chl_ug.cm.2, color = Genotype, shape = Cage_Uncaged), data = PAC_model_data) +
  labs(x = "Low Tide Mean Nitrate + Nitrite (umolL)", 
       y = expression(paste("Chlorophyll density (μg ", cm^2, ")")),
      title = " ") +
  scale_color_fish_d(option = "Coryphaena_hippurus", direction = -1) +
  theme_minimal()+
  scale_y_continuous(limits = c(0,4.5)) +
  theme(legend.position="none",
        plot.background = element_rect(fill = "white")) +
  guides(color = "none", shape = "none") +
  geom_line(data = pred_chl_NN, aes(x = x, y = predicted), color = "#21908c") +
  geom_ribbon(data = pred_chl_NN, aes(x = x, ymin = conf.low, ymax = conf.high),
              fill = "#21908c", alpha = 0.2)+
  annotate("text", y = 0.5, x = max(pred_chl_NN$x, na.rm = TRUE),
         label = paste("p =", round(p_values["Low_Tide_Mean_NN_umolL"], 6),
                       "\nMarginal R² =", round(marginal_r2, 3),
                       "\nConditional R² =", round(conditional_r2, 3)),
         hjust = 1.1, vjust = 1, size = 3)

##ggsave("output/pac_chl_NN_mod1.png", plot_chl_NN, height = 5, width = 5.5)
```


```{r}
polynomial_chl_model_pac <- lmer(Chl_ug.cm.2 ~ poly(pc1, 2) +
                                  (1|Genotype), data = PAC_model_data)
#check_model(polynomial_growth_model_pru)
Anova(polynomial_chl_model_pac)
```

```{r}
pred_chl_poly <- ggpredict(polynomial_chl_model_pac, terms = "pc1")

#annnotations
anova_results <- car::Anova(polynomial_chl_model_pac)
p_values <- anova_results$`Pr(>Chisq)`[1]  # Adjust index based on your model's fixed effects
names(p_values) <- rownames(anova_results)[1] 

plot_chl <- ggplot() +
  # Raw data points
  geom_point(data = PAC_model_data, 
             aes(x = pc1, y = Chl_ug.cm.2), color = "grey50", alpha = 0.6) +
  # Prediction line with confidence interval
  geom_line(data = pred_chl, aes(x = x, y = predicted), color = "#21908c") +
  geom_ribbon(data = pred_chl, aes(x = x, ymin = conf.low, ymax = conf.high), 
              fill = "#21908c", alpha = 0.2) +
  scale_y_continuous(limits = c(-1.5,25)) +
  labs(
    x = "PC1",
    y = expression(paste("Chlorophyll density (μg ", cm^2, ")"))
  ) +
  custom_theme+
  annotate("text", y = 1, x = max(pred_chl$x, na.rm=TRUE), 
           label = paste("p =", round(p_values["poly(pc1, 2)"], 6)), 
           hjust = 1.1, vjust = 1, size = 3)
```

```{r}
linear_chl_pac_growth <- lmer(Percent_Change ~ Chl_ug.cm.2 +
                                  (1|Genotype), data = PAC_model_data)

plot_chl_growth <- ggplot() +
  # Raw data points
  geom_point(data = PAC_model_data, 
             aes(x = Chl_ug.cm.2, y = Percent_Change,
                 color= Genotype, shape = Cage_Uncaged)) +
  labs(
    y = "Percent Change in Buoyant Weight",
    x = expression(paste("Chlorophyll density (μg ", cm^2, ")"))
  ) +
  scale_color_fish_d(option = "Coryphaena_hippurus", direction = -1)+
  custom_theme+
  annotate("text", y = 1, x = 1, 
           label = "p =0.1516", 
           hjust = 1.1, vjust = 1, size = 3)



```


## Porites rus
```{r}
pru_chl_raw <- ggplot(PRU_model_data, aes(x = pc1, y = Chl_ug.cm.2, color = Genotype, shape = Cage_Uncaged)) +
  geom_point() +
  labs(x = "PC1", 
       y = expression(paste("Chlorophyll density (μg ", cm^2, ")")),
      title = " ") +
  scale_color_fish_d(option = "Coryphaena_hippurus", direction = -1) +
  theme_minimal()+
  theme(legend.position="none",
        plot.background = element_rect(fill = "white")) +
  guides(color = "none", shape = "none") 

pru_chl_raw

#ggsave("output/pru_chl_nutrientpca_raw.png",pru_chl_raw, height = 5, width = 5.5)
```

### linear
```{r}
linear_chl_model_pru <- lmer(Chl_ug.cm.2 ~ pc1 +
                                  (1|Genotype), data = PRU_model_data)
#check_model(linear_growth_model_pru)
Anova(linear_chl_model_pru)
```

### polynomial
```{r}
polynomial_chl_model_pru <- lmer(Chl_ug.cm.2 ~ poly(pc1, 2) +
                                  (1|Genotype), data = PRU_model_data)
#check_model(polynomial_growth_model_pru)
Anova(polynomial_chl_model_pru)
```